Learning Auditory Similarities Between Instrumental Playing Techniques

### 2020-05-15 paper/paper.tex
fix typo
$k\geq 1$

The previous section described our methods for extracting spectrotemporal modulations in audio signals, as well as learning a non-Euclidean similarity metric between them.
We now turn to apply these methods to the problem of allocating isolated musical notes to clusters of some timbre similarity graph $\mathcal{G}$.
In practice, for training purposes, the cluster graph $\mathcal{G}$ stems from the consensus of $K=31$ users of the Cyberlioz web application, which was described in the \lnameref{sec:survey} section ($\mathcal{G}=\mathcal{G}_0$).
However, for evaluation purposes, this cluster graph corrsponds to the subjective preferences of a single user $k$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
% browsing a large-scale digital library of isolated musical notes, involving an extended vocabulary of playing techniques.

@@

Lastly, we consider the impact of alternative choices for feature extraction.
Let us recall the full definition of joint time--frequency scattering:
\begin{equation}
\mathbf{S_2^{\mathrm{joint}}}\boldsymbol{x}(t,\lambda,\alpha,\beta) = 
\left(
\Big\vert
\big\vert
\boldsymbol{x} \overset{t}{\ast} \boldsymbol{\psi}_{\lambda}
\big\vert
\overset{t}{\ast}
\boldsymbol{\psi}_{\alpha}
\overset{\log_2 \lambda}{\ast}
\boldsymbol{\psi}_{\beta}
\Big\vert
\overset{t}{\ast} \boldsymbol{\phi}_{T}
\overset{\log \lambda}{\ast} \boldsymbol{\phi}_{F}\right)(t),
\label{eq:joint-scattering}
\end{equation}
where the symbol $\boldsymbol{\psi}_{\lambda}$ (resp. $\boldsymbol{\phi}_T$) denotes a Morlet wavelet of center frequency $\lambda$ (resp. a Gaussian low-pass filter of scale $T$).
Besides the \emph{joint} time--frequency scattering transform, the generative grammar of scattering transforms \cite{lostanlen2019chapter} also encompasses the \emph{separable} time--frequency scattering transform:
\begin{equation}
\mathbf{S_2^{\mathrm{sep}}}\boldsymbol{x}(t,\lambda,\alpha,\beta) = 
\left(
\bigg\vert
\Big\vert
\big\vert
\boldsymbol{x} \overset{t}{\ast} \boldsymbol{\psi}_{\lambda}
\big\vert
\overset{t}{\ast}
\boldsymbol{\psi}_{\alpha}
\Big\vert
\overset{t}{\ast} \boldsymbol{\phi}_{T}
\overset{\log_2 \lambda}{\ast}
\boldsymbol{\psi}_{\beta}
\bigg\vert
\overset{\log \lambda}{\ast} \boldsymbol{\phi}_{F}
\right)(t),
\label{eq:separable-scattering}
\end{equation}
which has been applied successfully to classify environmental sounds \cite{bauge2013icassp} as well as playing techniques involving periodic modulations \cite{wang2019ismir}.

@@

\subsection*{What is the baseline performance?}
Further simplifying the feature set by replacing the scattering coefficients with MFCCs gives a performance of $81.8\% \pm 7.2$.
Compared to the separable scattering (which includes both first- and second-order coefficients), we see that the MFCCs perform markedly worse.
In order to confirm that the additional information captured by the joint scattering is effectively important, another feature set is considered where the MFCCs are complemented with monomials in order to reach the dimensionality of the joint scattering.
To do so, the $40$ features $f_i$ of the MFCCs are multiplied together $f_{i, j} = f_i \times f_j, i \neq j$.
\ja{What are the $f_i$ here?
How are they defined?}%
% TODO VL: give reference to MFCCs.
The resulting feature set consists of the 40 MFCCs features $f_i$ and the $1140$ features $f_{i, j}$.
The resulting $ap@5$ of $81.5\% \pm 7.3$ indicates that the increase of dimensionality is has a small effect, confirming the relevance of the additional information provided by the joint scattering.

>>>>

The previous section described our methods for extracting spectrotemporal modulations in audio signals, as well as learning a non-Euclidean similarity metric between them.
We now turn to apply these methods to the problem of allocating isolated musical notes to clusters of some timbre similarity graph $\mathcal{G}$.
In practice, for training purposes, the cluster graph $\mathcal{G}$ stems from the consensus of $K=31$ users of the Cyberlioz web application, which was described in the \lnameref{sec:survey} section ($\mathcal{G}=\mathcal{G}_0$).
However, for evaluation purposes, this cluster graph corrsponds to the subjective preferences of a single user $k\geq 1$, in which case we take $\mathcal{G}=\mathcal{G}_k$.
% browsing a large-scale digital library of isolated musical notes, involving an extended vocabulary of playing techniques.

@@

Let us recall the full definition of joint time--frequency scattering:
\begin{equation}
\mathbf{S_2^{\mathrm{joint}}}\boldsymbol{x}(t,\lambda,\alpha,\beta) = 
\left(
\Big\vert
\big\vert
\boldsymbol{x} \overset{t}{\ast} \boldsymbol{\psi}_{\lambda}
\big\vert
\overset{t}{\ast}
\boldsymbol{\psi}_{\alpha}
\overset{\log_2 \lambda}{\ast}
\boldsymbol{\psi}_{\beta}
\Big\vert
\overset{t}{\ast} \boldsymbol{\phi}_{T}
\overset{\log \lambda}{\ast} \boldsymbol{\phi}_{F}\right)(t),
\label{eq:joint-scattering}
\end{equation}
where the symbol $\boldsymbol{\psi}_{\lambda}$ (resp. $\boldsymbol{\phi}_T$) denotes a Morlet wavelet of center frequency $\lambda$ (resp. a Gaussian low-pass filter of scale $T$).
Besides the \emph{joint} time--frequency scattering transform, the generative grammar of scattering transforms \cite{lostanlen2019chapter} also encompasses the \emph{separable} time--frequency scattering transform:
\begin{equation}
\mathbf{S_2^{\mathrm{sep}}}\boldsymbol{x}(t,\lambda,\alpha,\beta) = 
\left(
\bigg\vert
\Big\vert
\big\vert
\boldsymbol{x} \overset{t}{\ast} \boldsymbol{\psi}_{\lambda}
\big\vert
\overset{t}{\ast}
\boldsymbol{\psi}_{\alpha}
\Big\vert
\overset{t}{\ast} \boldsymbol{\phi}_{T}
\overset{\log_2 \lambda}{\ast}
\boldsymbol{\psi}_{\beta}
\bigg\vert
\overset{\log \lambda}{\ast} \boldsymbol{\phi}_{F}
\right)(t),
\label{eq:separable-scattering}
\end{equation}
which has been applied successfully to classify environmental sounds \cite{bauge2013icassp} as well as playing techniques involving periodic modulations \cite{wang2019ismir}.

@@

\subsection*{What is the baseline performance?}
Lastly, we consider the impact of alternative choices for feature extraction.

$81.8\% \pm 7.2$

Arguably, this figure is not directly comparable with our best performing system, since

To disprove this hypothesis, 
Further simplifying the feature set by replacing the scattering coefficients with MFCCs gives a performance of $81.8\% \pm 7.2$.
Compared to the separable scattering (which includes both first- and second-order coefficients), we see that the MFCCs perform markedly worse.
In order to confirm that the additional information captured by the joint scattering is effectively important, another feature set is considered where the MFCCs are complemented with monomials in order to reach the dimensionality of the joint scattering.
To do so, the $40$ features $f_i$ of the MFCCs are multiplied together $f_{i, j} = f_i \times f_j, i \neq j$.
\ja{What are the $f_i$ here?
How are they defined?}%
% TODO VL: give reference to MFCCs.
The resulting feature set consists of the 40 MFCCs features $f_i$ and the $1140$ features $f_{i, j}$.
The resulting $ap@5$ of $81.5\% \pm 7.3$ indicates that the increase of dimensionality is has a small effect, confirming the relevance of the additional information provided by the joint scattering.
